|

### 2020-03-03 analysis/paper/manuscript.Rmd
Add link to GitHub repo to paper

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013]. While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way. Traditional formats (e.g. slides) for presenting information at stakeholder meetings are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community [@perkel2018]. As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018].

@@

The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background"). Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Catch"), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data. For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)). The dashboard also includes pages focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity"). Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages). The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents. However, there are two key benefits of this approach over the modus operandi of producing static documents and slides.  First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e. both broader patterns and finer details in the data are accessible via zooming and tooltips).  Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides, although, we note that this point is not exclusive to the proposed approach (e.g. slide shows can easily be made using R markdown). Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations. The dashboard also makes it easy to share the results with colleagues and stakeholders as it is rendered into a self-contained html file. This allows others to independently scrutinize details of both data and the model that are typically only accessible to the analyst. Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model. For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model. These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussions left us with the impression that this tool helped the assessment biologists, academics, resource managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables. Feedback and personal testimony from multiple stakeholders who have used the dashboard reaffirmed our impression that the tool is helpful as there was general agreement that access to the details eases communication and relieve concerns that something is hidden.

@@

@@

Visualizations play an important role in communicating fisheries science to a wide range of stakeholders [@levontin2017] and here we posit that interactive tools extend their utility by providing a richer approach to communicating scientific information. However, we must acknowledge there are costs to adopting the use of interactive tools. The first and obvious one being that staff are required to have both the time and training to effectively develop these tools. Ideally, professional programmers and graphic designers would be hired to construct dashboards and other interactive info-graphics [@mcinerny2014], but fisheries management organizations rarely have the resources to fund such initiatives. Consequently, the process of communicating fisheries science tends to be the responsibility of fisheries scientists who lack formal training in these fields. In our three examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming and graphic design, but all had their formal training in field-based population ecology. While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g. JavaScript, CSS, etc.). Moreover, a growing user base means that most programming issues were readily addressed through simple internet searches. There are also a growing number of exemplars (Appendix A), beyond what we show here, to inspire the development of new interactive tools. We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents. It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.

@@

>>>>

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013]. While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way. Visualisation play a critical role in communicating this information to stakeholders [@levontin2017], however, traditional formats (e.g. slides) are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community [@perkel2018]. As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018].

@@

The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background"). Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Catch"), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data. For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)). The dashboard also includes pages focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity"). Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages). The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents. However, there are two key benefits of this approach over the modus operandi of producing static documents and slides.  First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e. both broader patterns and finer details in the data are accessible via zooming and tooltips).  Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides, although, we note that this point is not exclusive to the proposed approach (e.g. slide shows can easily be made using R markdown). Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations.

The dashboard also makes it easy to share the results with colleagues and stakeholders as it is rendered into a self-contained html file. This allows others to independently scrutinize details of both data and the model that are typically only accessible to the analyst. Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model. For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model. These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussions left us with the impression that this tool helped the assessment biologists, academics, resource managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables. Feedback and personal testimony from multiple stakeholders who have used the dashboard reaffirmed our impression that the tool is helpful as there was general agreement that access to the details eases communication and relieve concerns that something is hidden.

@@

# Generic vs. bespoke

The tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations.

@@

Visualizations play an important role in communicating fisheries science to a wide range of stakeholders [@levontin2017] and here we posit that interactive tools extend their utility by providing a richer approach to communicating scientific information. However, we must acknowledge there are costs to adopting the use of interactive tools. The first and obvious one being that staff are required to have both the time and training to effectively develop these tools. Ideally, professional programmers and graphic designers would be hired to construct dashboards and other interactive info-graphics [@mcinerny2014], but fisheries management organizations rarely have the resources to fund such initiatives. Consequently, the process of communicating fisheries science tends to be the responsibility of fisheries scientists who lack formal training in these fields. In our three examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming and graphic design, but all had their formal training in field-based population ecology. While some learning time was needed, the learning curve was relatively minor because there was no need to learn a new programming language (e.g. JavaScript, CSS, etc.). We can confirm that an R user can construct a basic dashboard within a day, using Supplement 1 as a guide, and a more custom dashboard in a matter of weeks. Moreover, a growing user base means that most programming issues can be readily addressed through simple internet searches. There are also a growing number of exemplars (Appendix A), beyond what we show here, to inspire the development of new interactive tools. Once over the initial learning curve, the time investment may not be much greater than one would invest into a standard presentation. We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents. It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.

@@

All code used to generate this paper, Supplement 1 ("Getting Started" dashboard) and Supplement 2 ("NCAM explorer" dashboard) are available at https://github.com/PaulRegular/interactive-stock-assessment.
